---
title: "Bayesian Covid - DSBA 6010"
output:
  html_document:
    df_print: paged
  html_notebook: default
  pdf_document: default
---

#### Setting environment and packages:
```{r}
setwd("C:/Users/Mitchell Gaming PC/OneDrive - Visual Risk IQ, LLC/School Files/Spring 2022/Bayesian Stats/Actual Semester Project/Datasources")
library(rethinking)
set_cmdstan_path("C:/Users/Mitchell Gaming PC/Documents/.cmdstanr/cmdstan-2.28.2/cmdstan-2.28.2")
library(dagitty)
library(readr)
```
#### Loading the data:
```{r}
data <- read_csv('First_Draft_Data.csv')
```
#### Drawing an Initial DAG
```{r echo=TRUE}
g<- dagitty('dag {
bb="0,0,1,1"
"Proximity to Other States" [latent,pos="0.551,0.143"]
"incident rate" [outcome,pos="0.847,0.370"]
"proportion of >18 at risk" [pos="0.555,0.629"]
"status of vaccine mandate" [exposure,pos="0.304,0.375"]
"Proximity to Other States" -> "incident rate"
"proportion of >18 at risk" -> "incident rate"
"proportion of >18 at risk" -> "status of vaccine mandate"
"status of vaccine mandate" -> "incident rate"
}')
plot(g)
```
#### Standardizing data and putting it into dat list:
```{r}
data$R <- standardize(data['At-risk adults, as a share of all adults ages 18 and older'])
data$I <- standardize(data['Incident_Rate'])
data$V <- ifelse(data['Any Mandate in Place?']=="Yes", 1 , ifelse(data['Any Mandate in Place?']=="Prohibited" , 3 ,  2))

dat <- list(
  R = data$R,
  I = data$I,
  V = as.integer(data$V))
```
### Justifying Priors
Outcome:
Normally distributed Incident Rate (Number of Covid-19 Cases per 1000 population)

#### At Risk Adults:
Justified by prior scientific study of at-risk individuals in the United States by possession of 1 or more chronic conditions. Found that ~50% of Americans suffer, according to (https://www.cdc.gov/pcd/issues/2020/20_0130.htm#T1_down)

#### Vaccine Mandates
Index variable encoding:

1: 'No' Mandate.

2: 'Prohibited' legally from creating a mandate in this state

3: 'Yes' that a mandate has been enacted

### Total Effect of Vaccine Mandate
#### Initial Model Fit to Data (Will Sample Priors from Here)
```{r}
m1 <- ulam(
  alist(
    I ~ dnorm( mu , sigma ),
    mu <- a[V],
    a[V] ~ dnorm( 0 , 1 ),
    sigma ~ dexp(1)
  ), data=dat , chains=4 , log_lik=TRUE )
```
### Prior Predictive Check
```{r echo=TRUE}
prior <- extract.prior(m1)
prior$yes<-prior$a[,1]
prior$no<-prior$a[,2]
prior$prohibited<-prior$a[,3]
dens(prior$yes, lwd=4, col="blue", xlab="Total Effect of Mandate", xlim = c(-3, 3))
abline( v=0, lty=3)
dens(prior$no, lwd=4, col="red", add = TRUE)
dens(prior$prohibited, lwd=4, col="dark green", add = TRUE)
legend("topleft",
       c("Yes","No","Prohibited"),
       fill=c("blue","red","dark green"))
```
##### Implication of this prior:
Width: We hypothesized that the incidence would have some (moderate degree of) variability whether they did, did not have, or prohibited a mandate.

Index: The hypothesis illustrated here indicates that mandates would be associated with a similar distribution of incident rates. i.e., "Mandates don't do much to help"

Question: Could we set different priors for each level of the index variable?

#### Diagnostics: trace and rank plots
```{r echo=TRUE}
traceplot(m1)

trankplot(m1)

precis(m1, depth = 2)

postcheck(m1)
```

#### Visualizing Total Effect of Mandate
```{r echo=TRUE}
post1 <- extract.samples(m1)
post1$yes<-post1$a[,1]
post1$no<-post1$a[,2]
post1$prohibited<-post1$a[,3]
dens(post1$yes, lwd=4, col="blue", xlab="Total Effect of Mandate", xlim = c(-1.5, 1.5))
abline( v=0, lty=3)
dens(post1$no, lwd=4, col="red", add = TRUE)
dens(post1$prohibited, lwd=4, col="dark green", add = TRUE)
legend("topleft",
       c("Yes","No","Prohibited"),
       fill=c("blue","red","dark green"))
```

### Direct Effect of Mandate on Incident Rate
To do this, we must control for the characteristics of the population, since it is a confounder.

#### Creating the model
```{r}
m2 <- ulam( 
  alist(
    I ~ dnorm( mu , sigma ),
    mu <- a[V] + bR*R,
    a[V] ~ dnorm( 0 , 1 ),
    bR ~ dlnorm( 0 , 1),
    sigma ~ dexp(1)
  ), data=dat , chains=4)
```

#### Visualizing Priors
```{r}
prior <- extract.prior(m2)
plot( NULL , xlim=c(0,1) , ylim=c(0.5,1.5) ,
      xlab="At-Risk Population" , ylab="Incident Rate" )
title('Prior for Direct Effect of Mandate')
risk_seq <- seq( from=-0.1 , to=1.1 , length.out=30 )
vaccine_seq <-  rep(1, 30)
mu <- link( m2 , post=prior , data= data.frame(R=risk_seq, V = vaccine_seq))
for ( i in 1:50 ) lines( risk_seq , mu[i,] , col=col.alpha("black",0.3) )
```

We decided that these priors were too drastic of a relationship, and so we altered our priors.

```{r}
m2.1 <- ulam( 
  alist(
    I ~ dnorm( mu , sigma ),
    mu <- a[V] + bR*R,
    a[V] ~ dnorm( 0 , 1 ),
    bR ~ dlnorm( 0 , .5),
    sigma ~ dexp(1)
  ), data=dat , chains=4)
```

```{r}
prior <- extract.prior(m2.1)
plot( NULL , xlim=c(-2,2) , ylim=c(-2,2) ,
      xlab="At-Risk Population" , ylab="Incident Rate" )
title('Priors for Direct Effect of Mandate (v2)')
risk_seq <- seq( from=-2.5 , to=2.5 , length.out=30 )
vaccine_seq <-  rep(1, 30)
mu <- link( m2.1 , post=prior , data= data.frame(R=risk_seq, V = vaccine_seq))
for ( i in 1:50 ) lines( risk_seq , mu[i,] , col=col.alpha("black",0.3) )
```
These look much better.

#### Fit Diagnostics
```{r}
traceplot(m2.1)

trankplot(m2.1)

precis(m2.1, depth = 2)

postcheck(m2.1)
```

#### Plotting Posterior
All together:
```{r}
# Plotting Posterior all together
risk_seq <- seq( from=-2 , to=2 , length.out=30 )

plot( data[which(data$V == 1),]$R , data[which(data$V == 1),]$I , pch=16 , col='red' ,
      xlab="At-Risk Population" , ylab="Incident Rate" ,
      xlim=c(-2,2), ylim = c(-3,3))

points(data[which(data$V == 2),]$R , data[which(data$V == 2),]$I , pch=16 , col='dark green' )

points(data[which(data$V == 3),]$R , data[which(data$V == 3),]$I , pch=16 , col='blue' )

title('Incident Rate for Mandate Levels')

# Draw Yes Lines
mu_yes <- link( m2.1 , data=data.frame( V= 3, R=risk_seq ) )
mu_yes_mean <- apply( mu , 2 , mean )
lines( risk_seq , mu_yes_mean , lwd=2, col = 'blue')
#mu__yes_ci <- apply( mu , 2 , PI , prob=0.97 )
#shade( mu_yes_ci , risk_seq , col=col.alpha(rangi2,0.3) )

# Draw No Line
mu_no <- link( m2.1 , data=data.frame( V=1 , R=risk_seq ) )
mu_no_mean <- apply( mu_no , 2 , mean )
lines( risk_seq , mu_no_mean , lwd=2, col = 'red' )

# Draw Prohibited Line
mu_prohibited <- link( m2.1 , data=data.frame( V=2 , R=risk_seq ) )
mu_prohibited_mean <- apply( mu_prohibited , 2 , mean )
lines( risk_seq , mu_prohibited_mean , lwd=2, col = 'dark green')

# Draw Legend
legend("topleft",
       c("Yes","No","Prohibited"),
       fill=c("blue","red","dark green"))
```

Individual Levels:
```{r}
## Plot Individual Posteriors in Separate Graphs
risk_seq <- seq( from=-2 , to=2 , length.out=30 )
plot( data[which(data$V == 1),]$R , data[which(data$V == 1),]$I , pch=16 , col='black' ,
      xlab="At-Risk Population" , ylab="Incident Rate" ,
      xlim=c(-2,2), ylim = c(-3,3))
title('Incident Rate for Mandate Level = No')
lines( risk_seq , mu_no_mean , lwd=2, col = 'red' )
legend("topleft",
       c("No"),
       fill=c("red"))

plot( data[which(data$V == 2),]$R , data[which(data$V == 2),]$I , pch=16 , col='black' ,
      xlab="At-Risk Population" , ylab="Incident Rate" ,
      xlim=c(-2,2), ylim = c(-3,3))
title('Incident Rate for Mandate Level = Prohibited')
lines( risk_seq , mu_prohibited_mean , lwd=2, col = 'dark green')
legend("topleft",
       c("Prohibited"),
       fill=c("dark green"))

plot( data[which(data$V == 3),]$R , data[which(data$V == 3),]$I , pch=16 , col='black' ,
      xlab="At-Risk Population" , ylab="Incident Rate" ,
      xlim=c(-2,2), ylim = c(-3,3))
title('Incident Rate for Mandate Level = Yes')
lines( risk_seq , mu_yes_mean , lwd=2, col = 'blue')
legend("topleft",
       c("Yes"),
       fill=c("blue"))
```

